### 2020-01-24 draft.tex
Minor changes to draft wording, table labels, and legend placement on RIN charts

\title{{\Large \bf Industry Compliance Costs Under the Renewable Fuel Standard: Evidence from Compliance Credits\thanks{This paper has benefited from comments provided by William Shughart, Ben Blau, Megan Hansen, and two anonymous referees. The Center for Growth and Opportunity at Utah State University provided financial support for data procurement and research. All errors are the author's own.}}}
\author{Arthur R. Wardle\footnote{
PhD Student, Department of Agricultural \& Resource Economics, University of California, Berkeley. \href{mailto:arw@berkeley.edu}{\tt arw@berkeley.ed} (corresponding author).}\\
Sherzod B. Akhundjanov\footnote{Assistant Professor, Department of Applied Economics, Utah State University. \href{matilto:sherzod.akhundjanov@usu.edu}{\tt sherzod.akhundjanov@usu.edu}}}

@@

The Renewable Fuel Standard (RFS), which requires oil refineries to blend ethanol into domestic fuel supplies, is a market-based policy that implements tradable compliance credits (called RINs) to better equalize compliance costs across firms. We use variation in the prices of these credits to retrieve reduced-form estimates of how the RFS impacts the stock prices of refining firms. We take advantage of unanticipated regulatory announcements that caused major swings in RIN prices to implement an event study around those announcement dates and also fit bivariate time series models to firm $\times$ RIN pairs. Both models show no significant stock price response among small or medium size firms and a small but statistically significant price response among large refiners. These findings are relevant to policy in that they cast doubt on concerns that the RFS allows integrated refiners to abuse merchant refiners. Our findings also shed light on the necessity of Small Refinery Exemptions, which are intended to shield small, financially vulnerable refiners from RFS compliance costs.
\newline

@@

The Renewable Fuel Standard (RFS), created under the Energy Policy Act of 2005 and greatly expanded by the Energy Independence and Security Act of 2007, mandates the use of various biofuels in domestic transportation fuels.\footnote{A complete description of how the RFS works is available in \cite{Schnepf2013}.} The statute itself includes volumetric mandates for cellulosic, biomass-based biodiesel, ``advanced," and total renewable fuels through 2022. Obligated parties (fuel refiners and importers) are required to submit specified numbers of Renewable Identification Numbers (RINs) to comply with the standard. RINs are created by biorefineries, who generate them whenever they produce a gallon of biofuel and link them to those gallons. The RIN is then split from those gallons upon fuel blending and becomes available for use. RINs can be traded, allowing obligated parties to comply with RFS mandates either by blending renewable fuels themselves or buying excess RINs from other parties. The construction of the RIN market is functionally equivalent to other intensity standard credit trading schemes like Renewable Portfolio Standards and similar to other market-based environmental regulations such as pollution permits.

Because regulated firms can comply with the RFS either by blending additional biofuels or by buying RINs, the basic fundamental value of a RIN (with some complications, described later in this paper) is the marginal cost to the refining sector of blending an additional unit of biofuel over a unit of gasoline.

Even high RIN prices do not necessarily impact the bottom line of obligated parties if these costs are easy to pass through to consumers and demand for transportation fuels is inelastic. Indeed, a large literature establishes that refiners are able to fully (or even more than fully) pass RFS compliance costs onto consumers \citep{Knittel2017,Pouliot2017,Burkhardt2019,Li2019,Lade2019}. This, however, does not imply that the RFS has no financial impact on US oil refiners whatsoever. Complete pass-through does not replace the profit margins on refined crude oil that counterfactually would  have been sold without the mandate and may impose infrastructure costs to accommodate ethanol blends, negotiating costs with biorefineries, and myriad costs of doing business not fully captured in models of cost pass-through. So long as fuel supply and gasoline demand are not perfectly inelastic, higher mandates must cut gasoline's market share of fuel \citep{deGorter2009}. \cite{KnittelSmith2015} give a fuller description of ethanol's impact on oil refining profitability. The goal of our analysis is to establish how changes in the prices of RFS compliance credits (RINs) impact the value of the policy's obligated parties (refiners). To accomplish this, we implement two different reduced-form methods, which allow us to avoid imposing any particular causal channel for how the policy might impact firms' value.

@@

Outside of the academic pass-through literature, some stakeholders in RFS debates express concerns that larger refiners use the RFS to disadvantage smaller ones. Refiners in the United States can be split into ``merchant" refiners, who do not blend their own fuel and are generally smaller, and ``integrated" refiners, who do. Though refineries are the parties obligated by the RFS, RINs are not actually separated from biofuels until they are blended with gasoline. All this can be seen in Figure \ref{supplychain}, with integrated refineries owning both ``Oil Refinery" and ``Fuel Blender" assets and merchant refiners owning only the former.\footnote{Note that refining firms can also own biorefineries, as Valero does.}  Not owning blending assets leaves merchant refiners in the position of having to buy RINs on the market rather than being able to generate them themselves. Merchant refiners and their advocacy organizations often claim that being unable to generate RINs puts their operations at a competitive disadvantage and allows integrated refineries to sell excess RINs for windfall profits \citep[see discussion and footnotes in][p. 21-31]{EnvironmentalProtectionAgency2017}. The Environmental Protection Agency has dismissed such arguments, pointing primarily to economic research on RIN pass-through in doing so \citep{EnvironmentalProtectionAgency2017}. Research by \cite{Babcock2016} provides a theoretical explanation for why the RFS should not impact merchant and integrated refiners differentially.

@@

We directly examine the impact of RIN price fluctuations on the stock prices of obligated refineries. First, following \cite{Lade2018a}, who used event studies to examine how RIN price shocks affect commodity markets and biorefinery stocks, we use unanticipated regulatory announcements that drastically affected the price of RINs to identify impacts on every firm in our sample, as well as selected subgroups of firms. The two large shocks are plausibly exogenous and large in magnitude, allowing us to examine a direct response in the stock price of regulated firms. Second, we fit bivariate time series models for every possible firm $\times$ RIN combination. Modeling each firm separately allows us to investigate heterogeneity among firms. While these models provide a picture of how RINs and firm stock prices are associated over time, they are subject to endogeneity problems---both RINs and refining stocks are structurally related to commodity prices, fuel demand factors, and other variables. The intent of this paper is not to identify these structural relationships, and building them into the model is beyond the scope of this research. Instead, we use these models to validate and motivate our event study models, recognizing that endogeneity issues inescapably prevent us from causally interpreting the time series results, as well as to provide a graphical idea of how shocks affect firm value using impulse response functions.

@@

Understanding pass-through does get us most of the way to understanding how the RFS affects refiners financially, but pass-through is not the only avenue by which the RFS could impose compliance costs. Profit margins on gasoline sales lost to ethanol and costs associated with biofuel procurement are just two additional potential ways the RFS could impose costs on refiners, even at the margin of the current mandate and with complete pass-through. If positive shocks to the RIN market are unanticipated by investors, stock prices should adjust post-shock to reflect investor's revised expectations for future profits. The methods undertaken in this paper impose no particular cost channel on examining the RFS's impact, instead electing to allow financial markets price these anticipated costs. Examining stock price responses to variations in RIN prices gives a highly reduced-form but generalized and unstructured understanding of how the RFS impacts refineries.

@@

All sorts of heterogeneities in firm operations are of interest to regulatory agencies because differences in cost structures could counter-intuitively incite lower-cost firms to support the law as a new, artificial source of comparative advantage \citep{Salop1983}. Our reduced-form estimates of how firms respond to positive price changes in RFS compliance costs can lend credence to or discredit some of these theories.

@@

Unlike 2013, 2015's proposed and final rules, which were released on May 29 and November 30, are clearly and unambiguously apparent by visual inspection of Figure \ref{2015pricehistory}.\footnote{In an online supplement, \cite{Lade2018a} repeat the portion of their analysis detailing how policy announcements affect RIN prices for 2015 event dates. Unlike their 2013 analysis, they do not examine how 2015 announcements affected biofuel stocks or commodity markets.} The mandate modifications in the May 29, 2015 proposed rule were lower than expected; the price of D6 RINs dropped from \$0.69 one week before to \$0.3775 one week afterwards, a drop of more than 45\%. The final rule pushed the mandate back up, and a jump of similar magnitude to the original plummet surrounded that rule's announcement on November 30, 2015.

@@

For completeness, we also use multivariate time series methods to quantify the response of refining firm values to variations in RIN price series. In particular, for every Firm $\times$ RIN combination, we estimate impulse response functions using vector auto-regression models. These allow us to determine the dynamic response of the value of refining firms to shocks in the value of RIN prices. This portion of our analysis builds on past literature using time series methods to unravel how RINs transmit to wholesale and consumer fuels \citep{Knittel2017} and biofuel-related commodities \citep{Whistance2014, Whistance2016}. We perceive these models as being a secondary gut-check to the headline event study results.

@@

The mandate for D3 RINs is consistently minuscule, and markets for those RINs are commensurately quite thin. Therefore, we restrict our analysis to D6, D5, and D4 RINs. RIN series also differ by their year of creation; apart from some flexibility allowing limited inter-year banking and borrowing of RIN stocks, RINs are mostly used to comply with the mandate for the year in which they were generated. Thus, we limit our analysis to RINs of a single compliance year.\footnote{The majority of the literature examining RIN prices do not restrict their analyses to a single compliance year. It is not always clear how this literature stitches together price series for RINs belonging to difference compliance years, especially when RINs for multiple compliance years are sold contemporaneously. One simple approach, taking the average of all traded prices for a given RIN variety, requires imposing the additional assumption that banking/borrowing constraints are non-binding.} As discussed in Section \ref{sec_2015shocks}, we analyze data from 2015. Because of some pre-year trading and the fact that RINs are not actually submitted for compliance until a few months after year end, our data does extend a little beyond a single calendar year.

@@

\subsection{Firm Characteristics}\label{sec_firmcharacter}
This study encompasses all publicly traded firms with at least 200,000 barrels per day of refining capacity as of January 1, 2015, plus Western Refining \citep{EnergyInformationAdministration2015}. Below this cutoff, fewer firms are publicly traded and the publicly traded ones tend to be companies whose main industry is not fuels (e.g. Delta Airlines). We use market capitalization as a relevant measure of firm size \citep{Fama1992}, sorting firms into one of three bins as seen in figure \ref{bins}. The cutoffs were chosen somewhat arbitrarily to result in balanced bins, but, as the next paragraph further explains, the cutoffs do run parallel to important qualitative differences between firms.

Firms characterized as large (market capitalizations in excess of \$100B) are also the US firms generally considered to be large, integrated refineries: BP, Shell, Chevron, and ExxonMobil. In mid-2015, an analysis released by Valero Energy Corporation argued that those firms were already separating RINs in excess of their own RFS obligations, yielding them ``windfall profits," because their sales of branded gasoline made up large percentages of their refinery production \citep{ValeroEnergyCorporation2016}. Patterns of price response among these energy firms could lend evidence for or against those conjectures. The line between small and medium refining firms also separates relatively smaller nationwide operations from merely regional companies.

@@

Both methods of our analysis require stationary data inputs, so we begin by testing for stationarity in our data. Tests for stationarity are known to be biased towards conclusions of non-stationarity in the presence of structural breaks \citep{Perron1989}. However, despite evidence indicating the presence of structural breaks in RIN price series \citep{Mason2016}, no prior research to our knowledge reports results for stationarity using tests designed to account for breaks, though \cite{Lade2018a} does consider alternate tests accounting for breaks in a footnote.

@@

\noindent where $y_t$ is the variable whose stationarity is being tested; $\Delta$ is the difference operator; $\tilde{S}_t=y_t-\tilde{\psi}_x-Z_t\tilde{\delta}$ for $t\ge2$; $\tilde{\delta}$ are coefficients recovered from a regression of $\Delta y_t$ on $\Delta Z_t$ (i.e., $\Delta y_t=\delta'\Delta Z_t+e_t$);  $\tilde{\psi}_x=y_1-Z_1\tilde{\delta}$; and $u_t$ satisfies the normality conditions defined by \citet[p. 336]{Phillips1988}. The test statistic for the unit root null hypothesis is the $t$-statistic on the parameter estimate for $\phi$, and significance is checked using critical values for an exogenous 2-break unit root test reported in \cite{Lee2003}.\footnote{\cite{Lee2003} only report critical values for tests with a sample size of 100, whereas our sample size is 357. This will result in the test being underpowered.}

For testing stationarity of our non-RIN series, the LS test conveniently exhibits nice size and power properties even when the actual data generating process contains no breaks. Results of the LS test are reported in Table \ref{stationarity} alongside the typical Augmented Dickey-Fuller (ADF) and Kwiatkowski-Phillips-Schmidt-Shin (KPSS) tests, which do not account for structural breaks. The three tests conclude unanimously that every RIN and stock price series is non-stationary.

\begin{table}[htbp] \centering 
	\caption{Summary Statistics and Stationarity Tests} 
	\label{stationarity} 
	\begin{tabular}{@{\extracolsep{5pt}} lcccccc} 
		\hline 
		\hline \\[-1.8ex] 
		& \multicolumn{3}{c}{Summary Statistics}  & \multicolumn{3}{c}{Stationarity Tests}\\
		\cmidrule(r){2-4} \cmidrule(r){5-7}
		& Mean & St. Dev & Obs.        & ADF & KPSS & LS \\ 
		\hline
		Null & 	   &       &    & Non-stationary & Stationary & Non-stationary \\
		\hline \\[-1.8ex] 
		\input{tables/stationarity_edited.tex}
		\hline
	\end{tabular} 
\begin{flushleft}
\scriptsize{Significance at alpha levels of 0.1, 0.05, and 0.01 are reported with *, **, and ***, respectively.}\\
\end{flushleft}
\end{table}

@@

Event study models effectively characterize short-run, one-off responses to large market shocks. Regulatory changes that precipitate massive swings in compliance credit values offer a canonical setting for such a market shock. Event studies of compliance credit shocks can be decomposed to study regulation's distributional effects, as in \cite{Bushnell2013}, who counterintuitively show that stock prices of carbon-intensive firms actually fell in response to a slump in EU carbon prices. In the context of the RFS, \cite{Lade2018a} apply event studies to 2013 shocks to examine the policy's effect on commodity markets and biorefining firms.

@@

\begin{table}[!htbp] \centering 
	\caption{Results from Event Studies} 
	\label{eventstudies} 
	\resizebox{.7\textwidth}{!}{
		\begin{tabular}{@{\extracolsep{5pt}} lcccc} 
			\hline 
			\hline \\[-1.8ex] 
			& Large Firms & Medium Firms & Small Firms & All Firms \\ 
			\hline \\[-3.6ex] 
			{\it \textbf{Event 1}} & & & & \\
			\input{tables/event_studies_edited.tex}
			\hline 
		\end{tabular} 
	}
	\begin{flushleft}
		\scriptsize{Note: Each column reports selected parameter estimates and significance levels for an estimation of Equation \ref{eventstudy} on a subset of firms. See Table \ref{bins} to examine the firms in each size bin. Hypothesis testing is conducted using the sample quantile test described in \cite{Gelbach2013}. Significance at alpha levels of 0.1, 0.05, and 0.01 are reported with *, **, and ***, respectively.}\\
	\end{flushleft}
\end{table}

The results from the event study analysis are reported in Table \ref{eventstudies}. As in \cite{Lade2018a}, p-values are based on sample quantile tests for significance designed to accommodate event studies with few firms, as discussed in \cite{Gelbach2013}.\footnote{More information on the test in the context of difference-in-differences modeling is available in \cite{Conley2011}.} The basic idea behind this test is to use the empirical cdf of non-event date residuals as the distribution for hypothesis testing, since event date parameters are identified off of variation within a single day.

The first event does not register for any firms in the event studies. From Figure \ref{2015pricehistory}, the first event date seems to affect D6 RINs most dramatically. The effects on D5 and D4 RINs, however, are smaller and seemingly opposite to one another. If biodiesel and advanced ethanol really are the marginal compliance fuels under the RFS, as \cite{Lade2018a} reveals and our bivariate time series results confirm, the patterns in the RINs following the first event should not be expected to generate significant responses in firm value.

In contrast, Figure \ref{2015pricehistory} shows that the second event has a much clearer and consistently positive price impact on all RINs. Firms with small and medium market capitalizations do not significantly react to the event, and that lack of response is reflected in the pooled regression as well. But firms with large market capitalizations do experience losses at lag 4, with nearly-significant losses continuing to lag 5. Lacking an exact causal channel, it is difficult to confidently explain the lag in these effects, though it should be noted that RIN prices did not make their full upward adjustment in a single day.

For robustness, we re-ran each of these event study models using the WTI oil price as our normal return control. The broad pattern of results stayed similar, though the significance level of large firms' fourth lag dropped to 10\% and the fifth lag dropped significance completely.

@@

Time series methods are commonly used to analyze RIN price behavior  \citep{Mason2016} and the relationships between RINs and other commodity or asset prices \citep{Serra2011, Whistance2014, Whistance2016, Knittel2017}. To corroborate our event study findings, we fit bivariate time series models for each RIN $\times$ Firm pair in our data. We initially test each pair for cointegation; finding none, we address non-stationary in this multivariate context by estimating vector autogressive models (VAR) in differences. Using these results, we visualize the effect of an exogeneous shock to RIN prices on firm values using impulse response functions. We discuss limitations of these models, and how to consider them in light of the event study results.

@@

To select an appropriate bivariate time series model for a pair of non-stationary series, we first consider the possibility that they are cointegrated \citep{Engle1987}. While cointegration tests are affected by structural breaks, structural breaks have little impact on the size or power of Johansen's cointegration test \citep{Campos1996}. For that reason, we elect to use Johansen's maximum eigenvalue and trace tests without modification \citep{Johansen1991} to evaluate the relationship between each pairwise combination of RIN series and firms. Table \ref{cointegration} reports the results of these tests.

@@

The tests for cointegration do no reveal any cointegrating relationships significant at the 5\% level among any of the firm $\times$ RIN pairs under consideration. A smattering of pairs demonstrate cointegration at a 10\% significance level; apart from HollyFrontier's potential cointegration with D6 RINs, these pairs are all large firms and biodiesel RINs.

>>>>

\title{{\Large \bf Industry Compliance Costs Under the Renewable Fuel Standard: Evidence from Compliance Credits\thanks{This paper has benefited from comments provided by William Shughart, Ben Blau, Megan Hansen, and two anonymous referees. The Center for Growth and Opportunity at Utah State University provided financial support for data procurement and research. All errors are the author's own.}}}
\author{Arthur R. Wardle\footnote{
PhD Student, Department of Agricultural \& Resource Economics, University of California, Berkeley. \href{mailto:arw@berkeley.edu}{\tt arw@berkeley.edu} (corresponding author).}\\
Sherzod B. Akhundjanov\footnote{Assistant Professor, Department of Applied Economics, Utah State University. \href{matilto:sherzod.akhundjanov@usu.edu}{\tt sherzod.akhundjanov@usu.edu}}}

@@

The Renewable Fuel Standard (RFS), which requires oil refineries to blend ethanol into domestic fuel supplies, is a market-based policy that implements tradable compliance credits (called RINs) to better equalize compliance costs across firms. We exploit unanticipated regulatory announcements that caused major swings in RIN prices to retrieve reduced-form estimates of how the RFS impacts stock prices of refining firms. Our analysis reveals no significant stock price response among small or medium size firms and a small but statistically significant price response among large refiners. These findings are relevant to policy in that they cast doubt on concerns that the RFS allows integrated refiners to abuse merchant refiners. Our findings also shed light on the necessity of Small Refinery Exemptions, which are intended to shield small, financially vulnerable refiners from RFS compliance costs.
\newline

@@

The Renewable Fuel Standard (RFS), created under the Energy Policy Act of 2005 and greatly expanded by the Energy Independence and Security Act of 2007, mandates the use of various biofuels in domestic transportation fuels.\footnote{A complete description of how the RFS works is available in \cite{Schnepf2013}.} The statute itself includes volumetric mandates for cellulosic, biomass-based biodiesel, ``advanced," and total renewable fuels through 2022. Obligated parties (oil refiners\footnote{We use the terms  ``oil refiners" and ``refiners" interchangeably in this study. We explicitly indicate when we refer to biorefining.} and importers) are required to submit specified numbers of Renewable Identification Numbers (RINs) to comply with the standard. RINs are created by biorefineries, who generate them whenever they produce a gallon of biofuel and link them to those gallons. The RIN is then split from those gallons upon fuel blending and becomes available for use. RINs can be traded, allowing obligated parties to comply with RFS mandates either by blending renewable fuels themselves or buying excess RINs from other parties. The construction of the RIN market is functionally equivalent to other intensity standard credit trading schemes like Renewable Portfolio Standards and similar to other market-based environmental regulations such as pollution permits.

Because regulated firms can comply with the RFS either by blending additional biofuels or by buying RINs, the basic fundamental value of a RIN (with some complications, described later in the paper) is the marginal cost to the refining sector of blending an additional unit of biofuel over a unit of gasoline.

Even high RIN prices do not necessarily impact the bottom line of obligated parties if these costs are easy to pass through to consumers and demand for transportation fuels is inelastic. Indeed, a large literature establishes that refiners are able to fully (or even more than fully) pass RFS compliance costs onto consumers \citep{Knittel2017,Pouliot2017,Burkhardt2019,Li2019,Lade2019}. This, however, does not imply that the RFS has no financial impact on US oil refiners whatsoever. Complete pass-through does not replace the profit margins on refined crude oil that counterfactually would  have been sold without the mandate and may impose infrastructure costs to accommodate ethanol blends, negotiating costs with biorefineries, and myriad costs of doing business not fully captured in models of cost pass-through. So long as fuel supply and gasoline demand are not perfectly inelastic, higher mandates must cut gasoline's market share of fuel \citep{deGorter2009}.\footnote{\cite{KnittelSmith2015} give a fuller description of ethanol's impact on oil refining profitability.} The goal of our analysis is to establish how changes in the prices of RFS compliance credits (RINs) impact the value of the policy's obligated parties (refiners). To accomplish this, we implement two different reduced-form methods, which allow us to avoid imposing any particular causal channel for how the policy might impact firms' value.

@@

Outside of the academic pass-through literature, some stakeholders in RFS debates express concerns that larger refiners use the RFS to disadvantage smaller ones. Refiners in the United States can be split into ``merchant" refiners, who do not blend their own fuel and are generally smaller, and ``integrated" refiners, who do. Though oil refineries are the parties obligated by the RFS, RINs are not actually separated from biofuels until they are blended with gasoline. All this can be seen in Figure \ref{supplychain}, with integrated refineries owning both ``Oil Refinery" and ``Fuel Blender" assets and merchant refiners owning only the former.\footnote{Note that refining firms can also own biorefineries, as Valero does.}  Not owning blending assets leaves merchant refiners in the position of having to buy RINs on the market rather than being able to generate them themselves. Merchant refiners and their advocacy organizations often claim that being unable to generate RINs puts their operations at a competitive disadvantage and allows integrated refineries to sell excess RINs for windfall profits \citep[see discussion and footnotes in][p. 21-31]{EnvironmentalProtectionAgency2017}. The Environmental Protection Agency has dismissed such arguments, pointing primarily to economic research on RIN pass-through in doing so \citep{EnvironmentalProtectionAgency2017}. Research by \cite{Babcock2016} provides a theoretical explanation for why the RFS should not impact merchant and integrated refiners differentially.

@@

We directly examine the impact of RIN price fluctuations on the stock prices of obligated oil refineries. First, following \cite{Lade2018a}, who used event studies to examine how RIN price shocks in 2013 affected commodity markets and biorefinery stocks, we use unanticipated regulatory announcements from 2015 that drastically affected the price of RINs to identify impacts on every firm in our sample, as well as selected subgroups of firms. The two large shocks examined are plausibly exogenous and large in magnitude, allowing us to study a direct response in the stock price of regulated firms. Second, we fit bivariate time series models for every possible firm $\times$ RIN combination. Modeling each firm separately allows us to investigate heterogeneity among firms. While these models provide a picture of how RINs and firm stock prices are associated over time, they are subject to endogeneity problems---both RINs and refining stocks are structurally related to commodity prices, fuel demand factors, and other variables. The intent of this paper is not to identify these structural relationships, and building them into the model is beyond the scope of this research. Instead, we use these models to validate and motivate our event study models, recognizing that endogeneity issues inescapably prevent us from causally interpreting the time series results, as well as to provide a graphical idea of how shocks affect firm value using impulse response functions.

@@

Understanding pass-through does get us most of the way to understanding how the RFS affects refiners financially, but pass-through is not the only avenue by which the RFS could impose compliance costs. Profit margins on gasoline sales lost to ethanol and costs associated with biofuel procurement are just two additional potential ways the RFS could impose costs on refiners, even at the margin of the current mandate and with complete pass-through. If positive shocks to the RIN market are unanticipated by investors, stock prices should adjust post-shock to reflect investor's revised expectations for future profits. The methods undertaken in this paper impose no particular cost channel on examining the RFS's impact, instead electing to allow financial markets price these anticipated costs. Examining stock price responses to variations in RIN prices gives a reduced-form but generalized understanding of how the RFS impacts refineries.

@@

Heterogeneities in firm operations are of interest to regulatory agencies because differences in cost structures could counter-intuitively incite lower-cost firms to support the law as a new, artificial source of comparative advantage \citep{Salop1983}. Our reduced-form estimates of how different firms respond to positive price changes in RFS compliance costs can lend credence to or discredit some of these theories.

@@

Unlike 2013, 2015's proposed and final rules, which were released on May 29 and November 30 respectively, are clearly and unambiguously apparent by visual inspection of Figure \ref{2015pricehistory}.\footnote{In an online supplement, \cite{Lade2018a} repeat the portion of their analysis detailing how policy announcements affect RIN prices for 2015 event dates. Unlike their 2013 analysis, they do not examine how 2015 announcements affected biofuel stocks or commodity markets.} The mandate modifications in the May 29, 2015 proposed rule were lower than expected; the price of D6 RINs dropped from \$0.69 one week before to \$0.3775 one week afterwards, a drop of more than 45\%. The final rule pushed the mandate back up, and a jump of similar magnitude to the original plummet surrounded that rule's announcement on November 30, 2015.

@@

For completeness, we also use multivariate time series methods to quantify the response of refining firm values to variations in RIN price series. In particular, for every Firm $\times$ RIN combination, we estimate impulse response functions using vector autoregressive models. These allow us to determine the dynamic response of the value of refining firms to shocks in the value of RIN prices. This portion of our analysis builds on past literature using time series methods to unravel how RINs transmit to wholesale and consumer fuels \citep{Knittel2017} and biofuel-related commodities \citep{Whistance2014, Whistance2016}. We perceive these models as being a secondary analysis to the headline event study results.

@@

The mandate for D3 RINs is consistently minuscule, and markets for those RINs are commensurately quite thin. Therefore, we restrict our analysis to D6, D5, and D4 RINs. RIN series also differ by their year of creation; apart from some flexibility allowing limited inter-year banking and borrowing of RIN stocks, RINs are mostly used to comply with the mandate for the year in which they were generated. Thus, we limit our analysis to RINs of a single compliance year, similar to the analysis of \cite{Lade2018a}.\footnote{The majority of the literature examining RIN prices do not restrict their analyses to a single compliance year. It is not always clear how this literature stitches together price series for RINs belonging to different compliance years, especially when RINs for multiple compliance years are sold contemporaneously. One simple approach, taking the average of all traded prices for a given RIN variety, requires imposing the additional assumption that banking/borrowing constraints are non-binding.} As discussed in Section \ref{sec_2015shocks}, we analyze data from 2015. Because of some pre-year trading and the fact that RINs are not actually submitted for compliance until a few months after year end, our data does extend a little beyond a single calendar year.

@@

\subsection{Firm Characteristics}\label{sec_firmcharacter}
This study encompasses all publicly traded firms with at least 200,000 barrels per day of refining capacity as of January 1, 2015, plus Western Refining \citep{EnergyInformationAdministration2015}. Below this cutoff, fewer firms are publicly traded and the publicly traded ones tend to be companies whose main industry is not fuels (e.g., Delta Airlines). We use market capitalization as a relevant measure of firm size \citep{Fama1992}, sorting firms into one of three bins as seen in Table \ref{bins}. The cutoffs were chosen somewhat arbitrarily to result in reasonably balanced bins, but, as the next paragraph further explains, the cutoffs do run parallel to important qualitative differences between firms.

Firms characterized as large (market capitalizations in excess of \$100B) are also the US firms generally considered to be large, integrated refineries: BP, Shell, Chevron, and ExxonMobil. In mid-2015, an analysis released by Valero Energy Corporation argued that those firms were already separating RINs in excess of their own RFS obligations, yielding them ``windfall profits," because their sales of branded gasoline made up large percentages of their refinery production \citep{ValeroEnergyCorporation2016}. Patterns of price response among these energy firms could lend evidence for or against those conjectures. The line between small (market capitalizations less than \$10B) and medium (market capitalizations between \$10B and \$100B) refining firms also separates relatively smaller nationwide operations from merely regional companies.

@@

Both methods of our analysis require stationary data inputs, so we begin by testing for stationarity in our data. Tests for stationarity are known to be biased towards conclusions of non-stationarity in the presence of structural breaks \citep{Perron1989}. However, despite evidence indicating the presence of structural breaks in RIN price series \citep{Mason2016}, no prior research to our knowledge reports results for stationarity using tests designed to account for breaks, with the exception of \cite{Lade2018a} who do consider alternate tests accounting for breaks in a footnote.

@@

\noindent where $y_t$ is the variable whose stationarity is being tested; $\Delta$ is the difference operator; $\tilde{S}_t=y_t-\tilde{\psi}_x-Z_t\tilde{\delta}$ for $t\ge2$; $\tilde{\delta}$ are coefficients recovered from a regression of $\Delta y_t$ on $\Delta Z_t$ (i.e., $\Delta y_t=\delta'\Delta Z_t+e_t$);  $\tilde{\psi}_x=y_1-Z_1\tilde{\delta}$; and $u_t$ satisfies the normality conditions defined by \citet[p. 336]{Phillips1988}. The test statistic for the unit root null hypothesis is the $t$-statistic on the parameter estimate for $\phi$, and significance is checked using critical values for an exogenous 2-break unit root test reported in \cite{Lee2003}.

For testing stationarity of our non-RIN series, the LS test conveniently exhibits nice size and power properties even when the actual data generating process contains no breaks. Results of the LS test are reported in Table \ref{stationarity} alongside the typical Augmented Dickey-Fuller (ADF) and Kwiatkowski-Phillips-Schmidt-Shin (KPSS) tests, which do not account for structural breaks. The three tests conclude unanimously that every RIN and stock price series is nonstationary. Consequently, we use the differences series in our empirical analysis.

\begin{table}[htbp] \centering 
	\caption{Summary Statistics and Stationarity Tests} 
	\label{stationarity} 
	\begin{tabular}{@{\extracolsep{5pt}} lcccccc} 
		\hline 
		\hline \\[-1.8ex] 
		& \multicolumn{3}{c}{Summary Statistics}  & \multicolumn{3}{c}{Stationarity Tests}\\
		\cmidrule(r){2-4} \cmidrule(r){5-7}
		& Mean & St. Dev & Obs.        & ADF & KPSS & LS \\ 
		\hline
		Null & 	   &       &    & Nonstationary & Stationary & Nonstationary \\
		\hline \\[-1.8ex] 
		\input{tables/stationarity_edited.tex}
		\hline
	\end{tabular} 
\begin{flushleft}
\scriptsize{Note: ADF stands for augmented Dickey-Fuller test; KPSS stands for Kwiatkowski-Phillips-Schmidt-Shin test; and LS stands for Lee and Strazicich test, which is the preferred test. Significance at alpha levels of 0.1, 0.05, and 0.01 are reported with *, **, and ***, respectively.}\\
\end{flushleft}
\end{table}

@@

Event study models effectively characterize short-run, one-off responses to large market shocks \citep{MacKinlay1997}. Regulatory changes that precipitate massive swings in compliance credit values offer a canonical setting for such a market shock. Event studies of compliance credit shocks can be decomposed to study regulation's distributional effects, as in \cite{Bushnell2013}, who counterintuitively show that stock prices of carbon-intensive firms actually fell in response to a slump in EU carbon prices. In the context of the RFS, \cite{Lade2018a} apply event studies to 2013 shocks to examine the policy's effect on commodity markets and biorefining firms.

@@

\begin{table}[!htbp] \centering 
	\caption{Results from Event Studies} 
	\label{eventstudies} 
	\resizebox{.7\textwidth}{!}{
		\begin{tabular}{@{\extracolsep{5pt}} lcccc} 
			\hline 
			\hline \\[-1.8ex] 
			& Large Firms & Medium Firms & Small Firms & All Firms \\ 
			\hline \\[-3.6ex] 
			{\it \textbf{Event 1}} & & & & \\
			\input{tables/event_studies_edited.tex}
			\hline 
		\end{tabular} 
	}
	\begin{flushleft}
		\scriptsize{Note: Each column reports selected parameter estimates and significance levels for an estimation of Equation \ref{eventstudy} on a subset of firms. See Table \ref{bins} to examine the firms in each size bin. Hypothesis testing is conducted using the sample quantile (SQ) test described in \cite{Gelbach2013}. Significance at alpha levels of 0.1, 0.05, and 0.01 are reported with *, **, and ***, respectively.}\\
	\end{flushleft}
\end{table}

The results from the event study analysis are reported in Table \ref{eventstudies}. P-values are based on sample quantile tests for significance designed to accommodate event studies with few firms, as discussed in \cite{Gelbach2013}.\footnote{More information on the test in the context of difference-in-differences modeling is available in \cite{Conley2011}.} The basic idea behind this test is to use the empirical cdf of non-event date residuals as the distribution for hypothesis testing, since event date parameters are identified off of variation within a single day.

The first event (the proposed rule) does not register for any firms in the event studies. From Figure \ref{2015pricehistory}, the first event date seems to affect D6 RINs most dramatically. The effects on D5 and D4 RINs, however, are smaller and seemingly opposite to one another. If biodiesel and advanced ethanol really are the marginal compliance fuels under the RFS, as \cite{Lade2018a} reveals and our bivariate time series results confirm, the patterns in the RINs following the first event should not be expected to generate significant responses in firm value.

In contrast, Figure \ref{2015pricehistory} shows that the second event (the final rule) has a much clearer and consistently positive price impact on all RINs. Firms with small and medium market capitalizations do not significantly react to the event, and that lack of response is reflected in the pooled regression as well. But firms with large market capitalizations do experience losses at lag 4 (significant at a 5 percent level) with losses continuing to lag 5 (significant at a 10 percent level). This shows that changes in the prices of RINs do affect the value of large obligated refineries. Lacking an exact causal channel, it is difficult to confidently explain the lag in these effects, though it should be noted that RIN prices did not make their full upward adjustment in a single day.

For robustness, we also ran each of these event study models using the WTI oil price as our normal return control. The broad pattern of results stayed qualitatively similar, with large firms' fourth lag remaining statistically significant at a 10 percent level.

@@

Time series methods are commonly used to analyze RIN price behavior  \citep{Mason2016} and the relationships between RINs and other commodity or asset prices \citep{Serra2011, Whistance2014, Whistance2016, Knittel2017}. To corroborate our event study findings, we fit bivariate time series models for each RIN $\times$ Firm pair in our data. We initially test each pair for cointegation; finding none, we estimate bivariate vector autogressive models (VAR) in differences. Using these results, we visualize the dynamic response of firm values to exogeneous shocks in RIN prices using impulse response functions. We discuss limitations of these models, and how to consider them in light of the event study results.

@@

To select an appropriate bivariate time series model for a pair of nonstationary series, we first consider the possibility that they might be cointegrated \citep{Engle1987}. While cointegration tests are affected by structural breaks, structural breaks have little impact on the size or power of Johansen's cointegration test \citep{Campos1996}. For that reason, we elect to use Johansen's maximum eigenvalue and trace tests without modification \citep{Johansen1991} to evaluate the relationship between each pairwise combination of RIN series and firms. Table \ref{cointegration} reports the results of these tests.

@@

The tests for cointegration do no reveal any cointegrating relationships significant at the 5\% level among any of the firm $\times$ RIN pairs under consideration. A smattering of pairs demonstrate cointegration at a 10\% significance level; apart from HollyFrontier's (HFC) potential cointegration with D6 RINs, these pairs are all large firms and biodiesel RINs.
